### 2021-03-24 coverletter.Rmd
Update coverletter.Rmd

---
output: pdf_document
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
---

![](images/uncgreensboro_DoB_h_pms-c.png){width="200"}

Dear *Editor*

We are excited to submit our manuscript entitled '*Emergent dual scaling of riverine biodiversity*' to *Proceedings of the National Academy Sciences*.

Despite the inherent complexity of nature, some ecological patterns follow simple mathematical rules with far-reaching implications for biodiversity conservation. Much attention has been paid to the ubiquitous power-law scaling between species richness and area, in which species richness increases with area in a decelerating way. In spatially complex ecosystems, however, there may be an overlooked scaling law of biodiversity because 'scale-invariant' complexity of fractal ecosystem structure drives ecological dynamics in space. Branching river networks are an outstanding example of such complex ecosystems, where geometric ecosystem structure (e.g., branching patterns) remains constant across a wide range of spatial scales.

In this research, we used theory and analysis of extensive fish data to show that the scale-invariant complexity of branching networks is an alternative dimension of ecosystem structure to dictate riverine biodiversity patterns. Theory and empirical analysis agreed that, **regardless of ecological traits of constituent species, riverine biodiversity exhibited a dual scaling law along the two orthogonal dimensions of ecosystem size and scale-invariant complexity.** The elegant consistency of theoretical and empirical results provided strong evidence that the dual scaling of biodiversity is a pervasive yet overlooked feature of complex ecosystems.

Our study distinguishes itself from previous studies in three important ways. (1) Our study is the first to show the *dual* scaling law of biodiversity. Notably, this pattern has emerged irrespective of ecological contexts or geographic regions, emphasizing the generality of our findings. (2) Our empirical analysis used *watershed ecosystems* as individual replicates of statistical analysis; our data set included 180 watersheds with 6605 sites across Japan and the US and this extensive data enabled us to perform the ecosystem-level analysis. Therefore, unlike most previous studies focused on local biodiversity patterns *within a single watershed*, this study provides crucial insights into how ecological communities are structured *across watersheds*. (3) The present study provides the foundation for complexity-based ecosystem management. While area-based ecosystem management spearheads global conservation efforts (e.g., protected areas), we are unaware of any conservation scheme that explicitly accounts for ecosystem complexity in the design. Hence, our research provides a novel perspective to spatial planning of protected areas or ecological restoration.

In addition to the significance described above, the connection between network complexity and biodiversity has a broad implication for a truly interdisciplinary audience, including scientists working on fluvial geomorphology, network theory, and biodiversity. Our study is also timely given the unprecedented rate of global biodiversity loss in freshwater ecosystems. Therefore, *Proceedings of the National Academy Sciences* is an ideal outlet for this manuscript.

We suggest the following leading scientists in the field as potential reviewers:

-   Dr. Florian Altermatt (EAWAG; florian.altermatt\@eawag.ch): Lead a number of projects on biodiversity in river networks

-   Dr. Emanuel Fronhofer (Université de Montpellier; emanuel.fronhofer\@umontpellier.fr): Works on theoretical spatial ecology with a focus on complex networks

-   Dr. Jonathan Moore (Simon Fraser University; jwmoore\@sfu.ca): Works on emergent ecological properties of river networks

-   Dr. Justin Yeakel (University of California Merced; jdyeakel\@gmail.com): Develops theoretical models for ecological communities, including species interactions in a spatial context

-   Dr. Jean Gibert (Duke University; jean.gibert\@duke.edu): Develops general theoretical models for ecological communities

-   Dr. Winsor Lowe (University of Montana; winsor.lowe\@umontana.edu): Lead a number of projects on animal movements in stream networks

This work is original and is not under consideration in any other journal. All authors agree with the content and with the submission for publication, and all appropriate ethical and other necessary approvals were obtained for this research. We look forward to your reply and hope that you find our article suitable for publication in your journal.

Sincerely,

Akira Terui, Ph.D. (corresponding author)

University of North Carolina at Greensboro