### 2021-06-29 title.tex
Create title.tex

%\begin{titlepage}
\begin{flushleft}
\LARGE{\textbf{Supplementary Information for: \newline Emergent dual scaling of riverine biodiversity}}\\
\vspace*{\baselineskip}
\large{Akira Terui$^{*,1}$, Seoghyun Kim$^1$, Christine L. Dolph$^2$, Taku Kadoya$^3$, Yusuke Miyazaki$^4$}\\
\vspace*{1\baselineskip}
\normalsize{*Corresponding author \newline Email: hanabi0111@gmail.com}\\
\vspace*{1\baselineskip}
\normalsize{
$^1$ Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA \newline
$^2$ Department of Ecology, Evolution and Behavior, University of Minnesota, St. Paul, MN 55108, USA \newline
$^3$ National Institute for Environmental Studies, Tsukuba, Ibaraki 305-8506, Japan \newline
$^4$ Shiraume Gakuen College, Kodaira, Tokyo 187-8570, Japan}\\
\vspace*{2\baselineskip}
\textbf{This PDF file includes:}
\begin{itemize}
\item Supplementary text
\item Tables S1 - S7
\item Figures S1 - S20
\item SI References
\end{itemize}
\end{flushleft}
% \end{titlepage}

\newpage