### 2021-07-02 table_kable.Rmd

---
title: "Table"
output:
  word_document:
    reference_docx: word_format.docx
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

source(here::here("code_empirical/analysis_rlm_bf.R"))
op = function(x, d = 2) sprintf(paste0("%1.", d, "f"), x)

x  <-  lapply(fit, function(x) {

  conf <- confint.default(x$model)
  ci_test <- conf[,1] * conf[,2]

  paste0(op(coef(summary(x$model))[,1]),
         " (",
         op(coef(summary(x$model))[,2]),
         ")",
         ifelse(ci_test < 0, "", "*"))
})

Parameter <- c("$\\xi_0$",
              "$\\xi_1$", 
              "$\\xi_2$",    
              "$\\xi_3$",                             
              "$\\xi_4$",                                         
              "$\\xi_5$",
              "$\\xi_6$",
              "$\\xi_7$")

Effect <- c("Intercept",
            "$log_{10}$ Watershed area", 
            "$log_{10}$ Branching probability",    
            "Region (Midwest vs. Hokkaido)",                             
            "Air temperature",                                         
            "Precipitation",
            "Logit fraction of agriculture",
            "Dam density")

output <- tibble::as_tibble(do.call(cbind, x)) %>% 
  rename("$\\alpha$ diversity" = alpha,
         "$\\beta$ diversity" = beta,
         "$\\gamma$ diversity" = gamma) %>% 
  mutate(Effect,
         Parameter) %>% 
  relocate(Parameter,
           Effect)

```

**Table 1** Estimated parameters of robust regression models explaining fish species richness in Hokkaido (Japan) and the Midwest (US). The standard errors (SE) are shown in parenthesis. Response variables were log-10 transformed. Climate and land use variables (air temperature, precipitation, logit fraction of agriculture, dam density) are deviations from the regional averages and were standardized to a mean of zero and a standard deviation of one prior to the analysis.

```{r table1}

knitr::kable(output)

```

Note: * parameter estimates whose 95% confidence intervals (approximately $\xi_k \pm 2SE$) do not include zero